---
title: "R code for MAG phylogenetic analysis and comparisons to GTDB"
author: "Mantas Sereika"
date: "2025"
output:
  html_document: default
  pdf_document: default
---


### Load dependencies
```{r, include=TRUE, message=FALSE}
library(stringr)
library(ape)
library(caper)



library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggnewscale)
library(ggpubr)
library(ggtree)
library(ggtreeExtra)
library(treeio)
library(rsvg)
library(grid)
```


### Load data
```{r, include=TRUE}
mags <- read.delim("../analysis/datasets/dataset_S3.tsv", sep="\t", header=T)
gtdb_meta <- read.delim("../data/GTDB/gtdb_metadata_r220.tsv", sep="\t", header=T)
```


### Plot presets
```{r, include=TRUE}
preset_gen <- list(theme_bw(),
                   theme(legend.position = "bottom",
                         legend.text=element_text(size=7),
                         legend.title = element_text(size=7),
                         axis.title.y = element_text(size = 7), 
                         axis.title.x = element_text(size = 7),
                         axis.text.y = element_text(size = 6),
                         axis.text.x = element_text(size = 6),
                         panel.grid.minor = element_line(size = 0.15),
                         panel.grid.major = element_line(size = 0.3)))
```


### Function for splitting GTDB ranks
```{r, include=TRUE}
gtdb_rank <- function(class,pattern,trim_left,trim_right) {
  rank <-  str_extract(class,pattern)
  rank <-  gsub(trim_left, "", rank)
  rank <-  gsub(trim_right, "", rank)
  return (rank) }
```


### Wrangle GTDB taxonomies
```{r, include=TRUE}
# MFD-LR MAGs
mags$domain <- gtdb_rank(mags$GTDB_taxonomy,"d__[a-zA-Z0-9 -_]+;p__","d__",";p__")
mags$phylum <- gtdb_rank(mags$GTDB_taxonomy,";p__[a-zA-Z0-9 -_]+;c__",";p__",";c__")
mags$class <- gtdb_rank(mags$GTDB_taxonomy,";c__[a-zA-Z0-9 -_]+;o__",";c__",";o__")
mags$order <- gtdb_rank(mags$GTDB_taxonomy,";o__[a-zA-Z0-9 -_]+;f__",";o__",";f__")
mags$family <- gtdb_rank(mags$GTDB_taxonomy,";f__[a-zA-Z0-9 -_]+;g__",";f__",";g__")
mags$genus <- gtdb_rank(mags$GTDB_taxonomy,";g__[a-zA-Z0-9 -_]+;s__",";g__",";s__")
mags$species <- gtdb_rank(mags$GTDB_taxonomy,";s__[a-zA-Z0-9 -_]+",";s__","")

mags$phylum[is.na(mags$phylum)] <- paste0("Unclassified")
mags$class[is.na(mags$class)] <- paste0("Unclassified")
mags$order[is.na(mags$order)] <- paste0("Unclassified")
mags$family[is.na(mags$family)] <- paste0("Unclassified")
mags$genus[is.na(mags$genus)] <- paste0("Unclassified")
mags$species[is.na(mags$species)] <- paste0("Unclassified")

# GTDB genomes
gtdb_meta$domain <- gtdb_rank(gtdb_meta$gtdb_taxonomy,"d__[a-zA-Z0-9 -_]+;p__","d__",";p__")
gtdb_meta$phylum <- gtdb_rank(gtdb_meta$gtdb_taxonomy,";p__[a-zA-Z0-9 -_]+;c__",";p__",";c__")
gtdb_meta$class <- gtdb_rank(gtdb_meta$gtdb_taxonomy,";c__[a-zA-Z0-9 -_]+;o__",";c__",";o__")
gtdb_meta$order <- gtdb_rank(gtdb_meta$gtdb_taxonomy,";o__[a-zA-Z0-9 -_]+;f__",";o__",";f__")
gtdb_meta$family <- gtdb_rank(gtdb_meta$gtdb_taxonomy,";f__[a-zA-Z0-9 -_]+;g__",";f__",";g__")
gtdb_meta$genus <- gtdb_rank(gtdb_meta$gtdb_taxonomy,";g__[a-zA-Z0-9 -_]+;s__",";g__",";s__")
gtdb_meta$species <- gtdb_rank(gtdb_meta$gtdb_taxonomy,";s__[a-zA-Z0-9 -_]+",";s__","")
```


### Count genomes per GTDB ranks
```{r, include=TRUE}
gtdb_count <- function(df,rank) {
  # All MAGs
  count_all <-  aggregate(df[,rank], by=list(df[,rank]), FUN=length)
  colnames(count_all)[2] <- "gtdb_species"
  # HQ MAGs
  count_hq <- aggregate(df[df$mimag_high_quality == "t",rank], by=list(df[df$mimag_high_quality == "t",rank]), FUN=length)
  colnames(count_hq)[2] <- "gtdb_species_hq"
  # rRNA
  count_rrna <- aggregate(df[df$ssu_count > 0,rank], by=list(df[df$ssu_count > 0,rank]), FUN=length)
  colnames(count_rrna)[2] <- "gtdb_species_rrna"
  # Combine
  count_all <- merge(count_all,count_hq, by="Group.1", all=T)
  count_all <- merge(count_all,count_rrna, by="Group.1", all=T)
  count_all[is.na(count_all)] <- 0
  return (count_all) }

count_phylum <- gtdb_count(gtdb_meta,"phylum")
count_class <- gtdb_count(gtdb_meta,"class")
count_order <- gtdb_count(gtdb_meta,"order")
count_family <- gtdb_count(gtdb_meta,"family")
count_genus <- gtdb_count(gtdb_meta,"genus")
```


### Compare genome counts in GTDB metadata vs MFD MAGs
```{r, include=TRUE}
gtdb_comp <- function(count,mags,rank) {
  count$MAG <- ifelse((count$Group.1 %in% mags[,rank]), "Y","N")
  count_tmp <- aggregate(mags[,rank], by=list(mags[,rank]), FUN=length)
  colnames(count_tmp)[2] <- "mfd_mags"
  count <- merge(count,count_tmp, by="Group.1", all=TRUE)
  count <- count[!is.na(count$gtdb_species),]
  count$mfd_mags[is.na(count$mfd_mags)] <- 0
  count$total_mags <- count$gtdb_species + count$mfd_mags
  count$gain <- round(count$total_mags/count$gtdb_species*100-100,3)
  count$gain2 <- round(count$total_mags/count$gtdb_species,3)
  return (count) }

mags_novel <- mags[mags$species == "Unclassified" & mags$rep_status == "TRUE",]

count_phylum <- gtdb_comp(count_phylum,mags_novel,"phylum")
count_class <- gtdb_comp(count_class,mags_novel,"class")
count_order <- gtdb_comp(count_order,mags_novel,"order")
count_family <- gtdb_comp(count_family,mags_novel,"family")
count_genus <- gtdb_comp(count_genus,mags_novel,"genus")
```


### Plot MAG presence in taxonomic groups
```{r, include=TRUE}
gtdb_plot <- function(count,label,dotsize) {
  count <- count[order(count$gtdb_species,decreasing=TRUE),]
  count$ID <- seq.int(nrow(count))
  count_ <- count %>% dplyr::select(ID,gtdb_species,total_mags) %>%  pivot_longer(!ID, names_to = "group", values_to = "count")
  count_$group2 <- ifelse(count_$ID %in% count[count$mfd_mags == 0, ]$ID,"zero",count_$group)
  count_ <- count_[order(count_$group2,decreasing=TRUE),]
  
  plot <- ggplot(data=count_, aes(x=ID, y=count, color=group2)) + 
  geom_line(aes(group=ID),col="black", size=0.1, alpha =0.5) + geom_point(size=dotsize, alpha=0.5) + 
  labs(y=paste("Species per",label,"in GTDB",sep=" "), x="Taxa queue position by species count in decreasing order") +
  preset_gen + guides(color = guide_legend(override.aes = list(size = 1))) +
  scale_y_continuous(trans="log10", expand=c(0.05,0.05)) + 
  scale_x_continuous(expand=c(0.05,0.05), breaks = c(1, pretty(count$ID)[pretty(count$ID) != 0])) +
  scale_color_manual(name = "", values=c("gtdb_species"="#bf812d","total_mags"="#35978f", "zero"="#bdbdbd"),
                     labels=c("gtdb_species"="GTDB genomes","total_mags"="With MFD mags", "zero"="No MAGs recovered for group"))
  return (plot) }

plot_phylum <- gtdb_plot(count_phylum,"phylum",0.6)
plot_class <- gtdb_plot(count_class,"class",0.1)
plot_order <- gtdb_plot(count_order,"order",0.1)
plot_family <- gtdb_plot(count_family,"family",0.1)
plot_genus <- gtdb_plot(count_genus,"genus",0.1)

plot_cov <- ggarrange(plot_phylum,plot_class,plot_order,plot_family, labels = c("a","b","c","d"), font.label = list(size = 9),
                      nrow=4, ncol=1, align = "none", common.legend = TRUE, legend = "bottom")

#ggsave(plot_cov, file="../analysis/figures/FigS14.pdf", height = 240, width = 180, useDingbats=FALSE,  units ="mm", dpi = 300)
```


### Load phylogenetic tree
```{r, include=TRUE}
gtdb_tree <- read.tree("../data/GTDB/gtdb-r220_MFD-LR.tree")
```


### Wrangle tree and GTDB metadata
```{r, include=TRUE}
# Get species per phylum counts
gtdb_sub <- gtdb_meta[, c("accession","phylum")]
colnames(gtdb_sub) <- c("label","phylum")

gtdb_sub_count <- aggregate(gtdb_meta$species, by=list(gtdb_meta$domain,gtdb_meta$phylum), FUN=length)
colnames(gtdb_sub_count) <- c("domain",	"phylum",	"gtdb_species")

# Get MAG ids
tips <- as.data.frame(unique(gtdb_tree$tip.label))
mags_id <- as.data.frame(str_subset(tips[,1], "bin"))
colnames(mags_id) <- "label"

# Prune tree
gtdb_sub_ <- as.data.frame(gtdb_sub[, c("label")])
colnames(gtdb_sub_) <- "label"
sub <- as.data.frame(rbind(gtdb_sub_,mags_id))
gtdb_tree <- ape::keep.tip(gtdb_tree, c(as.character(sub$label)))

# Mark domains
tree_df <- gtdb_tree %>% as.treedata %>% as_tibble
gtdb_tree <- groupClade(gtdb_tree, .node=c(tree_df[(tree_df$label == "Prokaryotes"), ]$node,tree_df[(tree_df$label == "Archaea"), ]$node))

# Get data-frames with GTDB and MAG names and phyla classification
mags_ <- as.data.frame(mags_novel[, c("bin","phylum")])
colnames(mags_)[1] <- "label"
label_phyla <- as.data.frame(rbind(mags_,gtdb_sub))

# Make list of common phylum and label data-frame accordingly
abund_phyla <- as.data.frame(head(gtdb_sub_count[order(gtdb_sub_count$gtdb_species, decreasing = TRUE),]$phylum,15))
colnames(abund_phyla)[1] <- "phylum"

label_phyla$phylum_ <- ifelse((label_phyla$phylum %in% abund_phyla$phylum), label_phyla$phylum,"Other phyla")
list_phyla <- c(abund_phyla$phylum,"Other phyla")
```


### Plot tree
```{r, include=TRUE}
plot_tree <- ggtree(gtdb_tree, layout="circular", aes(color=group), lwd=0.05) %<+% label_phyla +
             scale_color_manual(values=c("#d8b365", "#74c476"), labels=c("Bacteria","Archaea")) +
             geom_point2(aes(subset=(label %in% mags$bin)), shape=21, size=1, stroke = 0, fill="#b10026") +  
             scale_fill_brewer(palette = "Paired")  +
             guides(color=guide_legend(order=1, ncol=2, title="Domain", override.aes = list(shape=NA, linewidth=1.5)), fill=none) + new_scale_fill() + 
             geom_fruit(geom=geom_tile, mapping = aes(y=label, x=1, fill=phylum_), width=0.2, pwidth=0.08, offset=-0.2) +
             guides(fill=guide_legend(order=2, ncol=2, title="Phylum", override.aes = list(size=2))) +
             theme(legend.text=element_text(size=7), legend.title = element_text(size=7)) +
             scale_fill_manual(breaks=list_phyla, values=c("#c777cb","#c9443c","#5b3687","#9cb241","#b2467e","#d1972c","#61bc6a","#bf702a",
                                                           "#45c097","#5e8bd5","#668433","#6d71d8","#c69e54","#ba4758","#ab5634","#d9d9d9"))

#ggsave(plot_tree, file="../analysis/figures/Fig4.pdf", height = 8, width = 12, useDingbats=FALSE)
ggsave(plot_tree, file="../analysis/figures/Fig4_.pdf", height = 120, width = 180, useDingbats=FALSE,  units ="mm", dpi = 300)
```


### Make alternative tree with all the branches from MAGs labelled
```{r, include=TRUE}
gtdb_tree2 <- gtdb_tree

tree_df2 <- gtdb_tree2 %>% as.treedata %>% as_tibble
tree_df2$mag <- ifelse(grepl("bin",tree_df2$label), 1, 0)
tree_df2$arc <- ifelse(tree_df2$label %in% gtdb_meta[gtdb_meta$domain=="Archaea",]$accession, 1, 0)
tree_df2$x <- ifelse(tree_df2$mag == 1, 0, ifelse(tree_df2$arc==1,2,1))

gtdb_tree2 <- groupOTU(gtdb_tree2, split(tree_df2$label, tree_df2$x))
tree_df2 <- gtdb_tree2 %>% as.treedata %>% as_tibble
```


### Plot alternative tree
```{r, include=TRUE}
plot_tree2 <- ggtree(gtdb_tree2, layout="circular", aes(color=group), lwd=0.05) %<+% label_phyla +
              scale_color_manual(values=c("#b10026","#d8b365", "#74c476"), labels=c("MFD MAGs","GTDB bacterial","GTDB archaeal")) +
              scale_fill_brewer(palette = "Paired")  +
              guides(color = guide_legend(order=1, ncol=3, title="MAG group:", override.aes = list(shape=NA, linewidth=2)), fill=none) + new_scale_fill() + 
              geom_fruit(geom = geom_tile, mapping = aes(y=label, x=1, fill=phylum_), width=0.2, pwidth=0.08, offset=-0.2) +
              guides(fill=guide_legend(order=2, title="Phylum", ncol=2)) +
              scale_fill_manual(breaks=list_phyla, values=c("#c777cb","#c9443c","#5b3687","#9cb241","#b2467e","#d1972c","#61bc6a","#bf702a",
                                                            "#45c097","#5e8bd5","#668433","#6d71d8","#c69e54","#ba4758","#ab5634","#d9d9d9")) 

ggsave(plot_tree2, file="../analysis/figures/FigS12_.pdf", height = 8, width = 12, useDingbats=FALSE)
```


### Combine plots from the phylogeny workflow
```{r, include=TRUE}
wf_arc <- rsvg("../data/GTDB/summary_arc.svg", width = 3000, height = 1500)
wf_bac <- rsvg("../data/GTDB/summary_bac.svg", width = 3000, height = 1500)

plot_wf_arc <- ggplot() + annotation_custom(rasterGrob(as.raster(wf_arc), interpolate = TRUE), 
                                            xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) + theme_void()

plot_wf_bac <- ggplot() + annotation_custom(rasterGrob(as.raster(wf_bac), interpolate = TRUE), 
                                            xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) + theme_void()

plot_wf <- ggarrange(plot_wf_bac,plot_wf_arc, nrow=2, ncol=1,common.legend = FALSE, labels = c("a","b"), font.label = list(size = 9))

#ggsave(plot_wf, file="../analysis/figures/FigS13.pdf", height = 180, width = 180, units ="mm", dpi=300, useDingbats=FALSE)
```

